Geometric variational problems

### 2020-11-12 NSF2020/Description/specificaim1_ver2.tex
Update specificaim1_ver2.tex
Fixed premature large deformation reference.

Hamm and Kozlov (HK) pioneered 
the modern theory of membrane continuum mechanics \cite{Hamm2000}, and their theory
%was pioneered by Hamm and Kozlov (HK) \cite{Hamm2000}. The theory 
is widely used to describe biological phenomona 
fission \cite{FrEsAkSh15, Maetal15, PhysRevE.79.031926},
fusion \cite{ChKo08, KoKo2002,Kuzmin7235,Aeffner2012},
poration \cite{Gaetal20}, phase boundaries, and interaction with inclusions
\cite{SeLeMaEg17,Saetal20, Pietal20}, which require resolution of the internal structure of the membrane.  
Recently, there has been a revival in interest in the HK theory as the quadratic assumption
for the elasticity energy density has caused
%Using the HK  model as a base,
%PI RR and collaborators calculated, for the first time in continuum theory, a least energy path for transitions between planar bilayers, a membrane stalk,
%hemifusion diaphragm and the fusion pore \cite{RyWaCo13,RyKlYaCo16},
%and the energies calculated by these continuum studies (Figure \ref{fig:barriers}A),
%are in agreement with barrier heights derived by molecular dynamics and experimental studies \cite{FrRoPi17}.  
%
\begin{wrapfigure}[11]{l}{0.47\textwidth}
\centerline{\includegraphics[width=0.47\textwidth]{Figures/Deformations.pdf}}
%  \centering{\includegraphics[width=0.80\textwidth]{Figures/Deformations.pdf}}
  \vspace{-5pt}
\caption{\label{fig:deformations} \footnotesize Sketch of the HK
 membrane model \cite{Hamm2000}.}
\end{wrapfigure}
%
researchers to question the applicability of \eqref{ansatz3} \todo{this
equation hasn't been presented at this point} for large curvatures
\cite{PhysRevLett.117.188102, ARGUDO20161619}. For example, Galimzyanov
{\em et al.}~\cite{C9SM02079A} countered that energies derived from
molecular dynamics and elasticity agree, even when curvatures are large.
This proposal will develop much-needed mathematical analysis to resolve
these controversies due to the assumptions in the HK theory.

@@

The HK framework assumes a three-dimensional lipid monolayers with internal structure
consisting of straight fibers that represent elongated hydrocarbon chains. HK assumed an
elastic energy density ${\cal W}$ that is quadratic in the Green-Lagrange strain tensor for membrane deformation of
this striated, internal structure.
%
%
%
%These fibers
%tilt and stretch with respect to the dividing surface--the surface formed lying between the hydrocarbon chains and polar heads of lipids.
%Accordingly, the deformation takes the form 
%\begin{equation}
%  \label{LMdeformation}
%{\bf x}(X_1, X_2, X_3) = {\bf x}_0(X_1, X_2) + \zeta(x_0(X_1, X_2), X_3) {\bf n}(x_0(X_1, X_2))
%\end{equation}
%where $(X_1,X_2,X_3)$ are points in a reference volume.
%The map $x_0$ parametrizes the dividing surface $\Sigma$, and the unit vector field ${\bf n}$ parametrizes the
%direction of the lipid tails in the deformed state \cite{doi:10.1021/jp075641w,KLAUDA20083074}.
%The parameter $X_3$ and the function $\zeta$ parametrizes the distance along the hydrocarbon chain in the reference state
%and the deformed state, respectively, and so $0 \leq X_3 \leq \delta$ where $\delta = $ 2.5 nm is a realistic monolayer thickness. 
%
%The elastic theory assumes an energy density that is quadratic in the Green-Lagrange strain tensor for ${\bf x}$. 
%The chains stretch and compress to satisfy incompressibility.
%Symmetries with respect to mirror reflection and in-plane rotation, the internal structure assumptions and incompressibility
%lead to an elastic surface energy over the dividing surface $\Sigma$. 
%
This energy decomposes into four fundamental and independent
deformations (Figure~\ref{fig:deformations}): splay ($\Div {\bf n}$),
twist ($\Curl {\bf n}$), saddle splay ($\det \nabla {\bf n}$), and tilt
${\bf t}={\bf n}/({\bf N}\cdot {\bf n}) - {\bf N}$ where ${\bf N}$ is
the unit surface normal:
\begin{equation}
\label{ansatz3}
{\cal W} \equiv \int_{\Sigma} 
  \tfrac{1}{2}\KB\left[ \left( \Div {\bf n} + k_0\right)^2 - k_0^2\right] 
+ \tfrac{1}{2}\KT (\Curl {\bf n})^2 + \KG  \det \nabla {\bf n} + \tfrac{1}{2}\KTH |{\bf t}|^2 \,dA.
\end{equation}
Here, 
%$(\nabla {\bf n})_{ij} = \partial_{x_j}{\bf n} \cdot {\bf e}_i$ is the surface gradient with
%orthonormal tangent vectors $\{{\bf e}_1, {\bf e}_2\}$.
%%\todo[inline]{YNY and BQ: What do you call a coordinate system where the metric is the identity at one point?}
%Also, the surface divergence 
%$\Div {\bf n}$ is the trace, and the surface curl $\Curl {\bf n}$ is the anti-trace (sum of the off-diagonal elements), of $\nabla {\bf n}$ respectively.
%respectively (Figure \ref{fig:distortions}A--C).
the deformations come with elastic coefficients: the bending modulus
$\KB$, twist modulus $\KT$, saddle-splay modulus $\KG$, and tilt modulus
$\KTH$.
%HK define the tilt vector ${\bf t} = {\bf n}/({\bf N}\cdot {\bf n}) - {\bf N}$ where ${\bf N}$ is the unit surface normal.
%
The parameter $k_0$ is the spontaneous curvature and determines the
preferred lipid splay~\cite{RoLi15,Kozlov2007}.  As an illustration, the
lipid DSPC has $k_0 = -0.1$ nm$^{-1}$ 
%as a result of  %having a relatively long acyl chain. 
and these lipids would line the  inner monolayer of a 
spherical liposome, because the addition of a positive splay $\Div \mathbf{n}$ in~\eqref{ansatz3}
to the negative spontaneous curvature $k_0$ leads to lower energy \cite{Kamal22245, C3SM51829A, RoLi15,FriedSeguin15}.
The subtraction of $k_0^2$ makes the energy distortion free
\cite{Helfrich73,PhysRevLett.113.248102,Hamm2000}.

Under spatial scales much larger than the
membrane thickness, membrane energy is well-characterized by the
Canham-Helfrich energy used throughout the fluid-structure literature
\cite{QiangDu09, Lowengrub07,KimLai2010_JCP, Hu, HuLaiSeolEtAl2016_JCP,
  qua-bir2014, qua-vee-you2019}.
%\todo[inline]{YNY: place your recent works here}
The Canham-Helfrich energy is actually a special case of
\eqref{ansatz3} obtained by setting ${\bf n} =  \pm {\bf N}$ (the $\pm$ depending on
orientation) and collapsing both monolayers onto the membrane midplane.
%There, splay $\Div n$ equal to twice the mean curvature, saddle splay
%$\det \nabla n$ equal to the Gauss curvature, and both twist and tilt
%are equal to zero.

>>>>

Hamm and Kozlov (HK) pioneered 
the modern theory of membrane continuum mechanics \cite{Hamm2000}, and their theory
%was pioneered by Hamm and Kozlov (HK) \cite{Hamm2000}. The theory 
is widely used to describe biological phenomona 
fission \cite{FrEsAkSh15, Maetal15, PhysRevE.79.031926},
fusion \cite{ChKo08, KoKo2002,Kuzmin7235,Aeffner2012},
poration \cite{Gaetal20}, phase boundaries, and interaction with inclusions
\cite{SeLeMaEg17,Saetal20, Pietal20}, which require resolution of the internal structure of the membrane.  
Recently, there has been a revival in interest in the HK theory as the quadratic assumption
for the elasticity energy density has caused
researchers to question the applicability of the theory 
for large curvatures
\cite{PhysRevLett.117.188102, ARGUDO20161619}.
This proposal will develop much-needed mathematical analysis to resolve
these controversies due to the assumptions in the HK theory. 
%Using the HK  model as a base,
%PI RR and collaborators calculated, for the first time in continuum theory, a least energy path for transitions between planar bilayers, a membrane stalk,
%hemifusion diaphragm and the fusion pore \cite{RyWaCo13,RyKlYaCo16},
%and the energies calculated by these continuum studies (Figure \ref{fig:barriers}A),
%are in agreement with barrier heights derived by molecular dynamics and experimental studies \cite{FrRoPi17}.  
%
\begin{wrapfigure}[11]{l}{0.47\textwidth}
\centerline{\includegraphics[width=0.47\textwidth]{Figures/Deformations.pdf}}
%  \centering{\includegraphics[width=0.80\textwidth]{Figures/Deformations.pdf}}
  \vspace{-5pt}
\caption{\label{fig:deformations} \footnotesize Sketch of the HK
 membrane model \cite{Hamm2000}.}
\end{wrapfigure}
%

@@

The HK framework assumes a three-dimensional lipid monolayers with internal structure
consisting of straight fibers that represent elongated hydrocarbon chains. HK assumed an
elastic energy density ${\cal W}$ that is quadratic in the Green-Lagrange strain tensor for membrane deformation of
this striated, internal structure.
%
%
%
%These fibers
%tilt and stretch with respect to the dividing surface--the surface formed lying between the hydrocarbon chains and polar heads of lipids.
%Accordingly, the deformation takes the form 
%\begin{equation}
%  \label{LMdeformation}
%{\bf x}(X_1, X_2, X_3) = {\bf x}_0(X_1, X_2) + \zeta(x_0(X_1, X_2), X_3) {\bf n}(x_0(X_1, X_2))
%\end{equation}
%where $(X_1,X_2,X_3)$ are points in a reference volume.
%The map $x_0$ parametrizes the dividing surface $\Sigma$, and the unit vector field ${\bf n}$ parametrizes the
%direction of the lipid tails in the deformed state \cite{doi:10.1021/jp075641w,KLAUDA20083074}.
%The parameter $X_3$ and the function $\zeta$ parametrizes the distance along the hydrocarbon chain in the reference state
%and the deformed state, respectively, and so $0 \leq X_3 \leq \delta$ where $\delta = $ 2.5 nm is a realistic monolayer thickness. 
%
%The elastic theory assumes an energy density that is quadratic in the Green-Lagrange strain tensor for ${\bf x}$. 
%The chains stretch and compress to satisfy incompressibility.
%Symmetries with respect to mirror reflection and in-plane rotation, the internal structure assumptions and incompressibility
%lead to an elastic surface energy over the dividing surface $\Sigma$. 
%
This energy decomposes into four fundamental and independent
deformations (Figure~\ref{fig:deformations}): splay ($\Div {\bf n}$),
twist ($\Curl {\bf n}$), saddle splay ($\det \nabla {\bf n}$), and tilt
${\bf t}={\bf n}/({\bf N}\cdot {\bf n}) - {\bf N}$ where ${\bf N}$ is
the unit surface normal:
\begin{equation}
\label{ansatz3}
{\cal W} \equiv \int_{\Sigma} 
  \tfrac{1}{2}\KB\left[ \left( \Div {\bf n} + k_0\right)^2 - k_0^2\right] 
+ \tfrac{1}{2}\KT (\Curl {\bf n})^2 + \KG  \det \nabla {\bf n} + \tfrac{1}{2}\KTH |{\bf t}|^2 \,dA.
\end{equation}
Here, 
%$(\nabla {\bf n})_{ij} = \partial_{x_j}{\bf n} \cdot {\bf e}_i$ is the surface gradient with
%orthonormal tangent vectors $\{{\bf e}_1, {\bf e}_2\}$.
%%\todo[inline]{YNY and BQ: What do you call a coordinate system where the metric is the identity at one point?}
%Also, the surface divergence 
%$\Div {\bf n}$ is the trace, and the surface curl $\Curl {\bf n}$ is the anti-trace (sum of the off-diagonal elements), of $\nabla {\bf n}$ respectively.
%respectively (Figure \ref{fig:distortions}A--C).
the deformations come with elastic coefficients: the bending modulus
$\KB$, twist modulus $\KT$, saddle-splay modulus $\KG$, and tilt modulus
$\KTH$.
%HK define the tilt vector ${\bf t} = {\bf n}/({\bf N}\cdot {\bf n}) - {\bf N}$ where ${\bf N}$ is the unit surface normal.
%
The parameter $k_0$ is the spontaneous curvature and determines the
preferred lipid splay~\cite{RoLi15,Kozlov2007}.  As an illustration, the
lipid DSPC has $k_0 = -0.1$ nm$^{-1}$ 
%as a result of  %having a relatively long acyl chain. 
and these lipids would preferentially line the upper monolayer in Figure \ref{fig:deformations}B  
because the addition of a positive splay 
to the negative spontaneous curvature leads to lower energy \cite{Kamal22245, C3SM51829A, RoLi15,FriedSeguin15}.
The subtraction of $k_0^2$ makes the energy distortion free
\cite{Helfrich73,PhysRevLett.113.248102,Hamm2000}.

Although the HK elastic theory assumes small deformations, 
Galimzyanov {\em et al.}~\cite{C9SM02079A} have show that energies derived from
molecular dynamics and \eqref{ansatz3} are in agreement, even when curvatures are large.
Under spatial scales much larger than the
membrane thickness, membrane energy is well-characterized by the
Canham-Helfrich energy used throughout the fluid-structure literature
\cite{QiangDu09, Lowengrub07,KimLai2010_JCP, Hu, HuLaiSeolEtAl2016_JCP,
  qua-bir2014, qua-vee-you2019}.
%\todo[inline]{YNY: place your recent works here}
The Canham-Helfrich energy is actually a special case of
\eqref{ansatz3} obtained by setting ${\bf n} =  \pm {\bf N}$ (the $\pm$ depending on
orientation) and collapsing both monolayers onto the membrane midplane.
%There, splay $\Div n$ equal to twice the mean curvature, saddle splay
%$\det \nabla n$ equal to the Gauss curvature, and both twist and tilt
%are equal to zero.
